---
title: "PredPol Data Sheet"
author: "Surya Mattu (The Markup), Dhruv Mehrotra (For Gizmodo Media Group)"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: "cosmo"
params:
  client: "default"
  client_name: "default"
  first_prediction_date: as.Date("1900-01-01")
  final_prediction_date: as.Date("1900-01-02")
  confirmed_start: as.Date("1900-01-01")
  confirmed_end: as.Date("1900-01-02")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(targets)
library(mapdeck)
library(leaflet)
library(glue)
library(kableExtra)


## Data
tar_load(bg_ranked_pop_bounds_bg_race)
tar_load(demo_prop_buckets_bg_race)
tar_load(demo_prop_buckets_bg_income)

client_data <- bg_ranked_pop_bounds_bg_race %>%
  filter(department == params$client) %>%
  select(geoid, pred_count, total_race_est, contains("est"), contains("_tb"), -contains("moe"), geoid, state, county, tract, block_group, pred_pop_rank)
min_bg <- filter(client_data, pred_count == min(pred_count))
max_bg <- filter(client_data, pred_count == max(pred_count))

smallest_bg <- filter(client_data, total_race_est == min(total_race_est))
largest_bg <- filter(client_data, total_race_est == max(total_race_est))


plot_demo <- function(bg_demo_prop_at_pred_rank, analysis_demo) {
  plot.grey_box <- c(
    "Most Targeted\nBlock Groups",
    "Median Targeted\nBlock Groups",
    "Least Targeted\nBlock Groups",
    "Jurisdiction Total"
  )
  names(plot.grey_box) <- c("in_max_tb", "in_med_tb", "in_min_tb", "total")

  race.labs <- list(
    title = glue("Races composition of most, median and least targeted block groups"),
    x = "Race",
    y = "Proportion of Block Group ",
    levels = c("White", "Asian", "Latino", "Black"),
    factor_names = c("white" = "White", "asian" = "Asian", "latino" = "Latino", "black" = "Black")
  )

  income.labs <- list(
    title = glue("Household income composition of most, median and least targeted block groups"),
    subtitle = "",
    y = "Percentage Of Block Groups",
    levels = c(
      "Greater than 200k",
      "Between 120 and 150k",
      "Between 75 and 100k",
      "Less than 45k"
    ),
    factor_names = c(
      "lt.45k" = "Less than 45k",
      "bw.75.100k" = " $75k - 100k",
      "bw.120.150k" = "$125k - 150k",
      "gt.200k" = "$200k and above"
    )
  )

  plot.labs <- list(
    race = race.labs,
    income = income.labs
  )

  legend.labs <- list(
    race = "",
    income = ""
  )

  analysis_demo_label <- legend.labs[[analysis_demo]]

  bg_demo_prop_at_pred_rank %>%
    filter(department == params$client) %>%
    pivot_longer(
      contains("__"),
      names_to = "demo__prank",
      values_to = "values"
    ) %>%
    separate(demo__prank, c("demo", "prank_filter"), sep = "__") %>%
    separate(demo, c("demo", "pop_est"), sep = "_") %>%
    group_by(demo, prank_filter) %>%
    summarise(
      name = params$client_name,
      min_value = min(values),
      min_pop_est = paste(pop_est[which(values == min(values))], collapse = ", "),
      plot_name = "dept_prank_demo"
    ) %>%
    mutate(
      demo = plot.labs[[analysis_demo]][["factor_names"]][demo],
      demo = as_factor(demo),
      demo = fct_relevel(demo, levels = plot.labs[[analysis_demo]][["levels"]]),
    ) %>%
    # filter(prank_filter == targeting_group) %>%
    remove_missing() %>%
    ggplot(aes(x = demo, y = min_value)) +
    geom_col(aes(fill = demo)) +
    scale_fill_discrete() +
    scale_y_continuous(labels = scales::percent) +
    # geom_text(aes(label = format(round(min_value, digits = 1)*100), digits=2, nsmall=1), size = 4, nudge_y = 0.02,  color = "black") +
    facet_grid(. ~ prank_filter, labeller = labeller(prank_filter = plot.grey_box)) +
    labs(
      # title = plot.labs[[analysis_demo]][["title"]],
      # subtitle = plot.labs[[analysis_demo]][["subtitle"]],
      x = plot.labs[[analysis_demo]]["x"],
      y = plot.labs[[analysis_demo]][["y"]],
      fill = plot.labs[[analysis_demo]]["x"]
    ) +
    theme(
      legend.position = "right",
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
}


display_table <- function(data) {
  race.labs <- list(
    levels = c("White", "Asian", "Latino", "Black"),
    factor_names = c("white" = "White", "asian" = "Asian", "latino" = "Latino", "black" = "Black")
  )

  income.labs <- list(
    levels = c(
      "Greater than 200k",
      "Between 120 and 150k",
      "Between 75 and 100k",
      "Less than 45k"
    ),
    factor_names = c(
      "lt.45k" = "Less than 45k",
      "bw.75.100k" = " $75k - 100k",
      "bw.120.150k" = "$125k - 150k",
      "gt.200k" = "$200k and above"
    )
  )

  plot.labs <- list(
    race = race.labs,
    income = income.labs
  )


  nicer_labels <- c(
    "lt.45k" = "Less than 45k",
    "bw.75.100k" = " $75k - 100k",
    "bw.120.150k" = "$120k - 150k",
    "gt.200k" = "$200k and above",
    "white" = "White",
    "asian" = "Asian",
    "latino" = "Latino",
    "black" = "Black",
    "in_max_tb" = "Most Targeted Block Groups",
    "in_med_tb" = "Median Targeted Block Groups",
    "in_min_tb" = "Least Targeted Block Groups",
    "total" =   "Jurisdiction Total"
  )

  df <- data %>%
    filter(department == params$client) %>%
    pivot_longer(
      contains("__"),
      names_to = "demo__prank",
      values_to = "values"
    ) %>%
    separate(demo__prank, c("demo", "prank_filter"), sep = "__") %>%
    separate(demo, c("demo", "pop_est"), sep = "_") %>%
    group_by(demo, prank_filter) %>%
    summarise(
      name = params$client_name,
      min_value = min(values),
      min_pop_est = paste(pop_est[which(values == min(values))], collapse = ", "),
      plot_name = "dept_prank_demo"
    ) %>%
    mutate(
      demo = nicer_labels[demo],
      prank_filter = nicer_labels[prank_filter],
      min_value = round(min_value * 100, digits = 1),
      # demo = as_factor(demo),
      # demo = fct_relevel(demo, levels = plot.labs[[analysis_demo]][["levels"]]),
    ) %>%
    remove_missing() %>%
    select(demo, prank_filter, min_value) %>%
    arrange(desc(prank_filter))


  col_order <- c("prank_filter", "demo", "min_value")
  df2 <- df[, col_order]
  # df[order(df$`Accuracy`, df$Kappa, decreasing=TRUE),]
  return(df2 %>% rename(
    `Targeting Level ` = prank_filter,
    `Demographic` = demo,
    `Proportion of Block Group pop.` = min_value
  ))
}

plot_tb_map <- function(df, df_col) {
  pal_min <- leaflet::colorFactor(
    "PuRd",
    domain = df_col
  )

  leaflet::leaflet(df) %>%
    leaflet::addProviderTiles(providers$CartoDB.VoyagerLabelsUnder) %>%
    leaflet::addPolygons(
      color = "#999", weight = 0.6, opacity = 1,
      fillColor = ~ pal_min(df_col), fillOpacity = 0.3,
    ) %>%
    leaflet::addLegend(
      pal = pal_min, values = ~df_col, opacity = 0.5,
      title = "Most Targeted", position = "topright"
    )
}

geoid_fmt <- function(x) format(x, nsmall = 0, big.mark = "to")
pred_fmt <- function(x) format(x, nsmall = 0, big.mark = ",")

conditional <- function(tb, demo, data) {
  tb_stat <- stat_demo_at_tb(data, demo, tb)
  jur_stat <- stat_demo_at_tb(data, demo, "total")
  if (near(tb_stat, jur_stat)) {
    "similar"
  }
  else if (tb_stat > jur_stat) {
    "greater"
  } else if (tb_stat < jur_stat) {
    "smaller"
  }
}

stat_demo_at_tb <- function(bg_demo_prop_at_pred_rank, demo_group, targeting_group, client = params$client) {
  row <- bg_demo_prop_at_pred_rank %>%
    filter(department == client) %>%
    pivot_longer(
      contains("__"),
      names_to = "demo__prank",
      values_to = "values"
    ) %>%
    separate(demo__prank, c("demo", "prank_filter"), sep = "__") %>%
    separate(demo, c("demo", "pop_est"), sep = "_") %>%
    group_by(demo, prank_filter) %>%
    summarise(
      name = params$client_name,
      min_value = min(values),
      min_pop_est = paste(pop_est[which(values == min(values))], collapse = ", "),
      plot_name = "dept_prank_demo"
    ) %>%
    filter(demo == demo_group & prank_filter == targeting_group) %>%
    remove_missing()
  row$min_value
}
```


```{css, echo = FALSE}

       h1, h2, h3, h4, h5, h6, .h1, .h2, .h3, .h4, .h5, .h6{
           font-weight:600;
           line-height:1.1;
           color:#242a49;
       }

       h1.title{
           font-size:45px;
           font-family: 'Times New Roman', Times, serif;
           font-weight:bold;
           border-bottom:1px solid #ff335f;
           padding-top: 50px;
           margin-bottom:40px;
       }

       h1{
           font-size:28px;
           font-family: Arial, Helvetica, sans-serif;
       }
       
       h1.title {
         margin-bottom: 10px;
       }
       
       h2{
           font-size:18px;
           text-transform:uppercase;
           padding-top:30px;
           padding-bottom:10px;
       }

       body{
           font-family:Arial, Helvetica, sans-serif;
           font-size:16px;
           line-height:1.6;
           color:#242a49;
           background-color: #f3f2f1;
       }

       .nav-tabs > li > a:hover {
           border-color: #ff335f;
        }
        
        .nav-tabs > li.active > a, .nav-tabs > li.active > a:hover, .nav-tabs > li.active > a:focus {
            color: #242a49;
            background-color: #f3f2f1;
            border: 1px solid #ff335f;
            border-bottom-color: rgb(255, 51, 95);
            border-bottom-color: transparent;
            cursor: default;
            font-weight: bold;
        }
        
        .nav-tabs {
            border-bottom: 1px solid #ff335f;
        }

        #TOC .tocify {
            border: 1px solid #ff335f;
            border-radius: 0px;
        }

        .list-group-item.active, .list-group-item.active:hover, .list-group-item.active:focus {
            z-index: 2;
            color: #ffffff;
            background-color: #ff335f;
            border-color: #ff335f;
        }

        .nav > li > a {
            color: #242a49;
        }

        a:link{color:#ff335f;}

        .table > thead > tr > th, .table > tbody > tr > th, .table > tfoot > tr > th, .table > thead > tr > td, .table > tbody > tr > td, .table > tfoot > tr > td {
            border-top: 1px solid #ff335f;
        }    

        .table > thead > tr > th {
            vertical-align: bottom;
            border-bottom: 1px solid #ff335f;
        }
       
       .author {
          font-family: arial;
          font-size: 16px;
          padding-bottom: 18px;
        }   
```
# `r params$client_name`

The Markup and Gizmodo have obtained and analyzed actual predictions for more than three dozen departments that used PredPol predictive policing software for at least six months between 2018 and 2020. This data sheet provides the findings from our _disparate impact analysis_ and _public housing analysis_  for `r params$client_name` To learn more about the project read, our [investigation](https://themarkup.org/prediction-bias/2021/12/02/crime-prediction-software-promised-to-be-free-of-biases-new-data-shows-it-perpetuates-them). For more details on how we did this analysis, read our [methodology](https://themarkup.org/show-your-work/2021/12/02/how-we-determined-crime-prediction-software-disproportionately-targeted-low-income-black-and-latino-neighborhoods).

## Findings {.tabset}

### Overview
 - Predpol’s algorithm relentlessly targeted the block groups in each jurisdiction that were  most heavily populated by people of color and the poor, particularly those containing public housing. The algorithm spared block groups with more White residents the same level of scrutiny.

 - The proportion of each jurisdiction’s Black and Latino residents was higher in the most-targeted block groups and lower in the least-targeted block groups compared to the jurisdiction overall. The opposite was true for the White population: The least-targeted block groups contained a higher proportion of White residents, and the most-targeted block groups contained a lower proportion. 

 - For the majority of jurisdictions in our data set (27 jurisdictions), a higher proportion of their low-income households lived in the block groups that were targeted the most. In some jurisdictions, all of their subsidized and public housing was located in block groups PredPol targeted more than the median.
 
 - These vast disparities were caused by the algorithm relentlessly predicting crime in the block groups in each jurisdiction that contained a higher proportion of the low-income residents and Black and Latino residents. They were the subject of crime predictions every shift, every day, and in multiple locations in the same block group.
 
 - We also analyzed arrest statistics by race from the FBI's Uniform Crime Reporting (UCR) Project for 29 of the agencies in our data that were in UCR. In 90 percent of them, per capita arrests were higher for Black people than White people—or any other racial group included in the dataset, mirroring the characteristics of the neighborhoods that the algorithm targeted.
 
 - We analyzed arrest data provided by 10 law enforcement agencies in our data and the rates of arrest in predicted areas remained the same whether PredPol predicted a crime that day or not.
 

### Race and Ethnicity

 Compared to `r params$client_name`, overall, the __most-targeted block groups__ had:
 
 - A __`r conditional("in_max_tb", "asian", demo_prop_buckets_bg_race)`__ proportion  of __Asians__  residents.
 - A __`r conditional("in_max_tb", "black", demo_prop_buckets_bg_race)`__ proportion  of __Black__  residents.
 - A __`r conditional("in_max_tb", "latino", demo_prop_buckets_bg_race)`__ proportion of __Latino__ residents.
 - A __`r conditional("in_max_tb", "white", demo_prop_buckets_bg_race)`__ proportion  of __White__ residents.
 
  Compared to `r params$client_name` overall, the __least-targeted block groups__ had:
 
 - A __`r conditional("in_min_tb", "asian", demo_prop_buckets_bg_race)`__ proportion  of __Asian__ residents.
 - A __`r conditional("in_min_tb", "black", demo_prop_buckets_bg_race)`__ proportion  of __Black__ residents.
 - A __`r conditional("in_min_tb", "latino", demo_prop_buckets_bg_race)`__ proportion of __Latino__ residents.
 - A __`r conditional("in_min_tb", "white", demo_prop_buckets_bg_race)`__ proportion  of __White__ residents.


```{r total_jur_race , message=FALSE, warning=FALSE, fig.align="center" }
plot_demo(demo_prop_buckets_bg_race, "race")
```

```{r, message=FALSE, warning=FALSE}
knitr:::kable(display_table(demo_prop_buckets_bg_race)) %>%
  kable_styling(bootstrap_options = c("hover"), full_width = TRUE)
```

### Household Income

Compared to `r params$client_name` overall, the __most-targeted block groups__ had:
 
 - A __`r conditional("in_max_tb", "lt.45k", demo_prop_buckets_bg_income)`__ proportion  of households that made __less than $45K__ a year.
 - A __`r conditional("in_max_tb", "bw.75.100k", demo_prop_buckets_bg_income)`__ proportion  of households that made between __between $75K and 100k__ a year.
 - A __`r conditional("in_max_tb", "bw.120.150k", demo_prop_buckets_bg_income)`__ proportion of households that made between  __$120k and 150K__ a year.
 - A __`r conditional("in_max_tb", "gt.200k", demo_prop_buckets_bg_income)`__ proportion  of households that made __$200K and above__ a year.
 

Compared to the `r params$client_name` overall, the __least-targeted block groups__ had:
 
 - A __`r conditional("in_min_tb", "lt.45k", demo_prop_buckets_bg_income)`__ proportion  of households that made __less than $45K__ a year.
 - A __`r conditional("in_min_tb", "bw.75.100k",demo_prop_buckets_bg_income)`__ proportion  of households that made between __$75K and 100K__ a year.
 - A __`r conditional("in_min_tb", "bw.120.150k", demo_prop_buckets_bg_income)`__ proportion of households that made between  __$120K and 150K__ a year.
 - A __`r conditional("in_min_tb", "gt.200k", demo_prop_buckets_bg_income)`__ proportion  of households that made __$200K and above__ a year.
 


```{r total_jur_income , message=FALSE, warning=FALSE, fig.align="center" }
plot_demo(demo_prop_buckets_bg_income, "income")
```


```{r, message=FALSE, warning=FALSE}
knitr:::kable(display_table(demo_prop_buckets_bg_income)) %>%
  kable_styling(bootstrap_options = c("hover"), full_width = TRUE)
```

 
### Public Housing

```{r housing}
housing <- function() {
  tar_load(housing_preds)
  data <- filter(housing_preds, department == params$client)
  if (nrow(data) == 1) {
    paste0(c(glue("In {params$client_name} {format(round(data$pct_mtb*100),  nsmall=0, big.mark=", ")} percent of public housing was on block groups the software targeted the most, "), glue("{format(round(data$pct_gt_p50*100),nsmall=0, big.mark=', ')} percent of public housing was on block groups the software targeted more than the median.")), collapse = "\n")
  } else {{ glue("We did not find housing data for {params$client_name}") }}
}

analysis_date_confirmation <- function() {
  if (params$client != "cpso") {
    glue("We confirmed these dates with the {params$client_name}, police department.")
  } else {
    {
      glue("")
    }[]
  }
}
```
`r housing()`

The table below provides how many predictions each block with public housing received. The final column tells us the percentage of days a block received predictions from PredPol's software between __`r params$first_prediction_date`__ and __`r params$final_prediction_date`__. `r analysis_date_confirmation()`
 
```{r housing_table}

tar_load(housing_block_map)
housing_block_map %>%
  filter(client == params$client & housing_count > 0) %>%
  select(client, state, county, tract, block_group, block, predictions, housing_count, percentage) %>%
  mutate(
    fips = str_c(str_pad(state, 2, pad = "0"), str_pad(county, 3, pad = "0"), str_pad(tract, 6, pad = "0"), block_group),
    percentage = percentage * 100
  ) %>%
  arrange(desc(percentage)) %>%
  select(fips, block, predictions, housing_count, percentage) %>%
  rename(
    `Census GEOID` = fips, Block = block, `Predictions` = predictions, `Num. Public Housing Units` = housing_count, `Pct. days w/ Predictions` = percentage
  ) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("hover"), full_width = TRUE)
```


## Maps {.tabset}

### Predictions {.tabset}

#### Density Map 

The map below aggregates all the predictions `r {params$client_name}`, received in our analysis window into a 2D grid. 
Each square of the grid represents an area approximately 500 ft. x 500 ft., the size of the PredPol prediction box.
The color represents the number of predictions that occurred within the square.
The more predictions, the darker the square. 


```{r pred_count_map, message=FALSE, warning=FALSE, fig.show="hold", out.width="100%"}
preds <- read_csv(glue("out/pred_count_by_dept/{params$client}.csv"))
clien_geojson <- sf::read_sf(glue("datasheets/geojson/{params$client}-geojson-with-data.json"))
client_analysis <- client_data %>%
  mutate(
    id = str_c(str_pad(state, 2, pad = "0"), str_pad(county, 3, pad = "0"), str_pad(tract, 6, pad = "0"), block_group)
  )
client_map_data <- left_join(clien_geojson, client_analysis, by = "id")



MAPBOX_API_KEY <- "pk.eyJ1IjoidGhlbWFya3VwIiwiYSI6ImNrd2t3NnpmeDF3ZG4ybnFiYWVtc3o1bWcifQ.nS8d56mUd0SdX_FFsrRPOA"
mapdeck(token = MAPBOX_API_KEY, style = mapdeck_style("satellite-streets"), zoom = 15) %>%
  add_path(
    data = client_map_data,
    layer_id = "path_layer",
    stroke_colour = "#FFFFCC99",
    stroke_width = 2,
    stroke_opacity = 0.5
  ) %>%
  add_grid(
    data = preds,
    lat = "lat",
    lon = "lon",
    colour_function = "mean",
    colour = "pred_count",
    cell_size = 300,
    colour_range = colourvalues::colour_values(1:6, palette = "ylorrd", alpha = 0.6),
    layer_id = glue("grid_layer_{params$client}"),
    extruded = FALSE,
    # legend = TRUE,
    focus_layer = FALSE
  )
```



```{r pred_count_legend}
preds_bins <- tibble(
  count = preds$pred_count,
  x_cut = cut(count, 6)
) %>%
  mutate(x_tmp = str_sub(x_cut, 2, -2)) %>%
  separate(x_tmp, c("min", "max"), sep = ",") %>%
  mutate_at(c("min", "max"), as.double) %>%
  mutate(min = if_else(min < 0, 0, min)) %>%
  select(min, max) %>%
  arrange(max) %>%
  unique() %>%
  mutate(`Prediction Count` = glue("{min} - {max}"))

if (nrow(preds_bins) == 5) {
  preds_bins$Description <- c("Least Predictions", "", "", "", "Most Predictions")
  color_val <- colourvalues::colour_values(1:5, palette = "ylorrd", alpha = 0.6)
  preds_bins %>%
    select(`Prediction Count`, Description) %>%
    kable() %>%
    kable_styling(bootstrap_options = c("condensed"), full_width = F, position = "right") %>%
    column_spec(1,
      color = "black",
      background = color_val[1:5]
    )
} else {
  preds_bins$Description <- c("Least Predictions", "", "", "", "", "Most Predictions")
  color_val <- colourvalues::colour_values(1:6, palette = "ylorrd", alpha = 0.6)
  preds_bins %>%
    select(`Prediction Count`, Description) %>%
    kable() %>%
    kable_styling(bootstrap_options = c("condensed"), full_width = F, position = "right") %>%
    column_spec(1,
      color = "black",
      background = color_val[1:6]
    )
}
```
_Sources: Markup, Predpol_

The grid drawn on this map provides an approximate aggregation of the prediction data. The actual prediction box in the reports provided to departments will vary from the ones shown above.


#### Choropleth

This map shows the predictions aggregated to the level of the Census block group.
Aggregating prediction data to the geographic area of a Census block group introduces additional complexity to the analysis, and hence this map should be interpreted with some caution.
See the limitations section of the [methodology](https://themarkup.org/show-your-work/2021/12/02/how-we-determined-crime-prediction-software-disproportionately-targeted-low-income-black-and-latino-neighborhoods) for more details.


```{r pred_map_bg, message=FALSE, warning=FALSE}

make_map <- function(df, df_col, color_bin_palette, title) {
  pal <- leaflet::colorBin(
    color_bin_palette,
    domain = df_col
  )

  leaflet::leaflet(df) %>%
    leaflet::addProviderTiles(providers$CartoDB.Voyager) %>%
    leaflet::addPolygons(
      color = "#999", weight = 0.6, opacity = 1,
      fillColor = ~ pal(df_col), fillOpacity = 0.5,
    ) %>%
    leaflet::addLegend(
      pal = pal, values = ~df_col, opacity = 0.7,
      title = title, position = "bottomright"
    )
}

make_map(client_map_data, client_map_data$pred_count, "YlOrRd", "Prediction Count")
```
_Source: Markup, Predpol_

### Race and Ethnicity{.tabset}

#### Black
```{r pred_map_black, message=FALSE, warning=FALSE}
make_map(client_map_data, client_map_data$black_alone, "Blues", "Black Population")
```

_Source: 2018 five-year ACS._

#### Latino
```{r pred_map_latino, message=FALSE, warning=FALSE}
make_map(client_map_data, client_map_data$latino_alone, "Oranges", "Latino Population")
```

_Source: 2018 five-year ACS._

#### White
```{r pred_map_white, message=FALSE, warning=FALSE}
make_map(client_map_data, client_map_data$white_alone, "Reds", "White Population")
```
_Source: 2018 five-year ACS._

### Household Income{.tabset}

#### Less than $45k
```{r pred_map_45, message=FALSE, warning=FALSE}
make_map(client_map_data, client_map_data$lt.45k_est, "BuGn", "Less than $45K")
```

_Source: 2018 five-year ACS._

#### $75k - $100k
```{r pred_map_75, message=FALSE, warning=FALSE}
make_map(client_map_data, client_map_data$bw.75.100k_est, "BuPu", "$75K - $100K")
```

_Source: 2018 five-year ACS._

#### $125k - $150k
```{r pred_map_125, message=FALSE, warning=FALSE}
make_map(client_map_data, client_map_data$bw.120.150k_est, "RdPu", "$125K - $150K")
```

_Source: 2018 five-year ACS._

#### $200k and above
```{r pred_map_200, message=FALSE, warning=FALSE}
make_map(client_map_data, client_map_data$gt.200k_est, "OrRd", "$200K and above")
```
_Source: 2018 five-year ACS._

## Methods {.tabset}

We analyzed the distribution of PredPol predictions for `r params$client_name` at the geographic level of a Census block group, which is a cluster of blocks with a population of between a few hundred to a few thousand people, generally. There are **`r length(unique(client_data$geoid))` block groups** in `r params$client_name`, the smallest block group had a population of approximately __`r smallest_bg$total_race_est[1]`__ and the largest had a population of approximately __`r format(largest_bg$total_race_est[1], nsmall=0, big.mark=",")`__. 
```{r cpso_helper}
cpso_hack <- function() {
  if (params$client == "cpso") {
    ""
  } else {
    glue(", we received confirmation that {params$client_name} department used the software between  **{params$confirmed_start}** and **{params$confirmed_end}**")
  }
}
```
In `r params$client_name`, we analyzed **`r format(sum(client_data$pred_count), nsmall=0, big.mark=",")` predictions** and used there locations to determine the block groups that were targeted the most, the median and the least. This data sheet presents the breakdown of the racial groups and household income ranges of the people who lived in those block groups. We also present the breakdowns for `r params$client_name` overall for comparison. The predictions we analyzed were between **`r params$first_prediction_date`** and **`r params$final_prediction_date`** `r cpso_hack()`.

For the race/ethnicity and income analyses, we merged 2018 five-Year American Community Survey data and prediction data and observed the makeup of block groups that were targeted above and below the median, those targeted the most and those targeted the least. For the sake of consistency in our analysis we only used demographic groups for which we had reliable population estimates for all the jurisdictions in our data set. These are: 

 - Racial Groups
    - Black
    - Asian
    - Latino
    - White
 - Household Income
    - Less than $45K
    - Between $75K-$100K
    - Between $125K-$150K
    - Greater than $200K  


## Definitions {.tabset}

We used the Census’ “designated place” boundaries as the boundaries for most jurisdictions. For Sheriff’s departments we confirmed the boundaries with the department.

We defined the most-targeted block groups as those in `r params$client_name` which
encompassed the highest five percent of predictions. We defined the median-targeted
block groups as the five percent around the median block group for
predictions. And we defined the least-targeted block groups as those with the
bottom five percent of predictions.

In some of the larger jurisdictions, more than five percent of block groups got zero predictions. In those cases, we chose the most populated block groups with no predictions for the five percent. Learn more about how we did this in our [methodology](https://themarkup.org/show-your-work/2021/12/02/how-we-determined-crime-prediction-software-disproportionately-targeted-low-income-black-and-latino-neighborhoods).

We identified public housing through HUD's online lookup tool available at [https://resources.hud.gov](https://resources.hud.gov)

## Data

The data used to generate this analysis can be found in our [GitHub](https://github.com/the-markup/investigation-prediction-bias) repository. It also contains the URLs for the [rest of the data sheets](https://github.com/the-markup/investigation-prediction-bias#data-sheets) from our analysis.